Community Detection Methods for the Generalized Random Dot Product Graph Model

### 2021-05-25 proposal/proposal.tex
updated proposal

Graph and network data have become increasingly widespread in various
fields including sociology, neuroscience, biostatistics, and computer
science. This has resulted in various challenges for researchers who
rely on traditional statistical and machine learning methods, many of
which are incompatible with graph data and instead require the data to
exist as feature vectors in Euclidean space. Such challenges often
involve clustering and community detection, which are often the goal of
graph data analysis. Common clustering methods typically involve
calculating some central or representative point for each cluster around
which the data belonging to that cluster lie (e.g., Lloyd's algorithm
for \(K\)-means clustering \cite{1056489}, Gaussian Mixture Models
\cite{doi:10.1198/016214502760047131}). Because these methods involve
computing summary statistics within each cluster, such as the sample
average, they cannot be applied directly to graphs, necessitating
methods for transforming the graph data into feature vectors.

One family of methods for unifying graph community detection with
traditional clustering techniques is Spectral Clustering
\cite{DBLP:journals/corr/abs-0711-0189}, which involves embedding the
graph into Euclidean space, followed by applying a popular clustering
algorithm such as \(K\)-means clustering. The Random Dot Product Graph
(RDPG) \cite{10.1007/978-3-540-77004-6_11} and Generalized Random Dot
Product Graph (GRDPG) \cite{rubindelanchy2017statistical} models take
this further by explicitly constructing generative models such that
latent positions in Euclidean space are used to induce graphs. A
community detection algorithm motivated by this may involve learning the
latent positions given an observed graph and then learning the community
labels given the latent positions.

The aim of our research is to develop consistent community detection
techniques under the RDPG and GRDPG frameworks. First, we explore
existing generative graph models with underlying community structures
that can be inferred by connecting these models to the RDPG or GRDPG.
Then we explore other latent structures or mixture distributions in the
latent space that induce graphs for which consistent community detection
is possible.

@@

Let \(G = (V, E)\) be an undirected, unweighted graph with no self-loops
and \(n\) vertices. Denote \(A \in \{0, 1\}^{n \times n}\) as the
adjacency matrix for \(G\) such that \(A_{ij} = 1\) if there exists an
edge between vertices \(i\) and \(j\) and \(A_{ij} = 0\) otherwise.
Because \(G\) is symmetric and contains no self-loops,
\(A_{ij} = A_{ji}\) and \(A_{ii} = 0\) for \(i, j \in [n]\). We further
restrict our analyses to independent Bernoulli graphs. Let
\(P \in [0, 1]^{n \times n}\) be a symmetric matrix of edge
probabilities. Graph \(G\) is sampled from \(P\) by drawing
\(A_{ij} \stackrel{\text{indep}}{\sim}\text{Bernoulli}(P_{ij})\) for
each \(1 \leq i < j \leq n\) (\(A_{ji} = A_{ij}\) and \(A_{ii} = 0\)).
Finally, we denote
\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}\)
as the sample \(x_1, ..., x_n \in \mathbb{R}^d\) and
\(z_1, ..., z_n \in [K]\) as their corresponding (hidden) labels.

@@

Generative models for independent Bernoulli graphs involve defining the
edge probability matrix \(P\) whose \(ij\)\textsuperscript{th} entry is
the probability of an edge between vertices \(i\) and \(j\) for each
\(i, j \in [n]\). In order to motivate community detection methods, we
restrict the generative model such that for each pair of vertices, the
probability of an edge between the vertices is conditioned on their
labels. One such model is the Stochastic Block Model (SBM)
\cite{doi:10.1080/0022250X.1971.9989788}: Given \(K\) communities and
each vertex belonging to one community, the SBM restricts \(P\) to
\(K (K + 1) / 2\) unique entries such that \(P_{ij} = p_{z_i z_j}\)
where \(p_{kl} = p_{lk}\) is the probability of an edge between each
vertex in community \(k\) to each vertex in community \(l\). The
homogeneous SBM further restricts \(P\) to two unique entries:
\(P_{ij} = p\) if \(z_i = z_j\) and \(P_{ij} = q\) otherwise. Several
generalizations of the SBM have been introduced since, including the
Degree Corrected Block Model (DCBM) and the Popularity Adjusted Block
Model (PABM). Like the SBM, these models involve imposing some
restriction on \(P\) based on community labels (and possibly other
factors).

@@

Like the SBM, DCBM, and PABM, the RDPG and GRDPG are generative models
for graphs involving an edge probability matrix \(P\). The RDPG starts
with points in latent space \(X \in \mathcal{X} \subset \mathbb{R}^d\)
such that \(\forall x, y \in \mathcal{X}\), \(x^\top y \in [0, 1]\).
\(P\) is then constructed as \(P = X X^\top\), and graph \(G\) with
adjacency matrix \(A\) is drawn from \(P\). We provide a more formal
definition of the RDPG and GRDPG below.

\begin{definition}[(Generalized) Random Dot Product Graph]
Let $X \in \mathbb{R}^{n \times d}$ be a collection of $n$ points in $\mathcal{X} \subset \mathbb{R}^d$ such that $\forall x, y \in \mathcal{X}$, $x^\top y.\in [0, 1]$. $G = (V, E)$ is a Random Dot Product Graph if its adjacency matrix $A$ is drawn such that $A_{ij} \sim \text{Bernoulli}(x_i^\top x_j)$ for $i < j$, with $A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$. If on the other hand $A_{ij} \sim \text{Bernoulli}(x_i^\top I_{p, q} x_j)$ where $I_{p, q} = \text{blockdiag}(I_p, -I_q)$ and $p + q = d$, then $A$ is the adjacency matrix of a Generalized Random Dot Product Graph. These are denoted by $A \sim \text{RDPG}(X)$ and $A \sim \text{GRDPG}_{p, q}(X)$ respectively.

In addition, let $F$ be a probability distribution with support $\mathcal{X}$, and $x_1, ..., x_n \stackrel{iid}{\sim} F$ with $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. If $A$ is drawn from $X$ as before, then $(A, X) \sim \text{RDPG}(F, n)$ or $(A, X) \sim \text{GRDPG}_{p, q}(F, n)$. 
\end{definition}

@@

It is straightforward to show that all Bernoulli graphs with positive
semidefinite \(P\) are special cases of the RDPG, which is a special
case of the GRDPG, and all graphs generated by \(P\) are special cases
of the GRDPG. This includes the SBM, DCBM, and PABM. In the following
example, we show that under the RDPG framework, the latent configuration
that induces the SBM takes on a very particular form.

\begin{example}[Connecting the SBM to the RDPG]
Let $G = (V, E)$ with adjacency matrix $A$ be sampled from the SBM with two communities for which the within-edge probabilities are $p$ and $q$ for communities 1 and 2 respective, and the between-community edge probability is $r$, with $p q > r^2$. Let community 1 have $n_1$ vertices and community 2 have $n_2$ vertices such that $n_1 + n_2 = n$. Without loss of generality, organize $P$ and $A$ such that the $kl^{th}$ block represents edges between communities $k$ and $l$. Then $P = \begin{bmatrix} P^{(11)} & P^{(12)} \\ P^{(21)} & P^{(22)} \end{bmatrix}$ where each block is a constant value, e.g., $P^{(11)}_{ij} = p$. One RDPG representation of this SBM is:

@@

We first define the PABM.

\begin{definition}[Popularity Adjusted Block Model]
\label{pabm}
Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
set of $n$ 
vertices, $V$. Each vertex has a community label $1, ..., K$, and the rows and 
columns of $P$ are arranged by community label such that $n_k \times n_l$ block 
$P^{(kl)}$ describes the edge probabilities between vertices in communities 
$k$ and $l$ ($P^{(lk)} = (P^{(kl)})^\top$). 
Let graph $G = (V, E)$ be an undirected, unweighted graph such 
that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
realization of $\text{Bernoulli}(P)$, i.e., 
$A_{ij} \stackrel{\text{indep}}{\sim}\text{Bernoulli}(P_{ij})$ for $i > j$ 
($A_{ij} = A_{ji}$ and $A_{ii} = 0$).

@@

It is straightforward to show that the PABM, as well as all graphs such
that the adjacency matrix is drawn such that
\(A_{ij} \stackrel{\text{indep}}{\sim}Bernoulli(P_{ij})\), is a special
case of the GRDPG. It can also be shown that the latent positions of the
PABM under the GRDPG framework consists of \(K\) \(K\)-dimensional
subspaces in \(\mathbb{R}^{K^2}\). While there is no unique latent
configuration \(X\) such that \(X X^\top = P\), the edge probability
\(P\) for the PABM, they all have this subspace structure, and one in
particular consists of \emph{orthogonal} subspaces.

@@

While direct maximization of the likelihood or log-likelihood is
NP-complete, it is possible to apply an Expectation Maximization (EM)
algorithm to this problem. To set this up,

\hypertarget{generalizations}{%
\section{Generalizations}\label{generalizations}}

@@

If we extend this example such that \(n_1\) points are sampled uniformly
from the segment \(f_1(t) = (t, 0)\) and \(n_2\) points are sampled
uniformly from the segment \(f_2(t) = (t, a)\) for \(t \in [0, 1]\),
then a sample of size \(\min(n_1, n_2) \geq N(\delta, a)\) is sufficient
to satisfy:

@@

We can then further extend this to the case where points are drawn from
unit segments with noise.

@@

@@

>>>>

Graph and network data have become increasingly widespread in various
fields including sociology, neuroscience, biostatistics, and computer
science. This has resulted in various challenges for researchers who
rely on traditional statistical and machine learning methods, many of
which are incompatible with graph data and instead require the data to
exist as feature vectors in Euclidean space. This includes graph
clustering and community detection, which are often the goal of data
analysis on graphs and networks. Common clustering methods typically
involve calculating some central or representative point for each
cluster around which the data belonging to that cluster lie (e.g.,
Lloyd's algorithm for \(K\)-means clustering \cite{1056489}, Gaussian
Mixture Models \cite{doi:10.1198/016214502760047131}). Because these
methods involve computing summary statistics within each cluster, such
as the sample average, they cannot be applied directly to graphs,
necessitating methods for transforming the graph data into feature
vectors.

One family of methods for unifying graph community detection with
traditional clustering techniques is Spectral Clustering
\cite{DBLP:journals/corr/abs-0711-0189}, which involves embedding the
graph into Euclidean space, followed by applying a popular clustering
algorithm such as \(K\)-means clustering. The Random Dot Product Graph
(RDPG) \cite{10.1007/978-3-540-77004-6_11} and Generalized Random Dot
Product Graph (GRDPG) \cite{rubindelanchy2017statistical} models take
this further by explicitly constructing generative models for graphs
involving latent vectors in Euclidean space. In such a model, each
vertex has a corresponding vector in latent space, and the edge
probabilities between pairs of vertices is computed by operations on
corresponding pairs of latent vectors. A community detection algorithm
motivated by this may involve learning the latent positions given an
observed graph and then learning the community labels given the latent
positions.

The aim of our research is to develop consistent community detection
techniques under the RDPG and GRDPG frameworks. First, we explore
existing generative graph models with underlying community structures
that can be inferred by connecting these models to the RDPG or GRDPG. In
particular, we use the connection between the Popularity Adjusted Block
Model \cite{307cbeb9b1be48299388437423d94bf1} and the GRDPG to motivate
two community detection methods. Then we explore other latent structures
or mixture distributions in the latent space that induce graphs for
which consistent community detection is possible.

@@

Let \(G = (V, E)\) be an undirected, unweighted graph with no self-loops
and \(n\) vertices. Denote \(A \in \{0, 1\}^{n \times n}\) as the
adjacency matrix for \(G\) such that \(A_{ij} = 1\) if there exists an
edge between vertices \(i\) and \(j\) and \(A_{ij} = 0\) otherwise.
Because \(G\) is symmetric and contains no self-loops,
\(A_{ij} = A_{ji}\) and \(A_{ii} = 0\) for \(i, j \in [n]\). We further
restrict our analyses to independent Bernoulli graphs. Let
\(P \in [0, 1]^{n \times n}\) be a symmetric matrix of edge
probabilities. Graph \(G\) is sampled from \(P\) by drawing
\(A_{ij} \stackrel{\text{ind}}{\sim}\text{Bernoulli}(P_{ij})\) for each
\(1 \leq i < j \leq n\) (\(A_{ji} = A_{ij}\) and \(A_{ii} = 0\)). We
denote \(A \sim \text{BernoulliGraph}(P)\) as graph with adjacency
matrix \(A\) sampled from edge probability matrix \(P\) in this manner.
If each vertex has a hidden label in \([K]\), they are denoted as
\(z_1, ..., z_n\). Finally, we denote
\(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}\)
as the sample \(x_1, ..., x_n \in \mathbb{R}^d\).

@@

Generative models for independent Bernoulli graphs involve defining the
edge probability matrix \(P\) whose \(ij\)\textsuperscript{th} entry is
the probability of an edge between vertices \(i\) and \(j\) for each
\(i, j \in [n]\). In order to motivate community detection methods, we
restrict the generative model such that for each pair of vertices
\((i, j)\), the probability of an edge between the vertices \(P_{ij}\)
depends in some way on their labels \(z_i\) and \(z_j\).

One such model is the Stochastic Block Model (SBM)
\cite{doi:10.1080/0022250X.1971.9989788}: Given \(K\) communities and
each vertex belonging to one community, the SBM restricts \(P\) to
\(K (K + 1) / 2\) unique entries such that \(P_{ij} = p_{z_i z_j}\)
where \(p_{kl} = p_{lk}\) is the probability of an edge between each
vertex in community \(k\) to each vertex in community \(l\). The
homogeneous SBM further restricts \(P\) to two unique entries:
\(P_{ij} = p\) if \(z_i = z_j\) and \(P_{ij} = q\) otherwise. Several
generalizations of the SBM have been introduced since, including the
Degree Corrected Block Model (DCBM) \cite{Karrer_2011} and the
Popularity Adjusted Block Model (PABM)
\cite{307cbeb9b1be48299388437423d94bf1}. Like the SBM, these models
involve imposing some restriction on \(P\) based on community labels
(and possibly other factors).

@@

Like the SBM, DCBM, and PABM, the RDPG and GRDPG are generative models
for graphs involving an edge probability matrix \(P\). The RDPG starts
with points in latent space \(X \in \mathcal{X} \subset \mathbb{R}^d\)
for \(\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top y \in [0, 1]\}\).
\(P\) is then constructed as \(P = X X^\top\), and
\(A \sim \text{BernoulliGraph}(P)\). We provide a more formal definition
of the RDPG and GRDPG as follows:

\begin{definition}[(Generalized) Random Dot Product Graph]
Let $X \in \mathbb{R}^{n \times d}$ be a collection of $n$ points in $\mathcal{X} \subset \mathbb{R}^d$ where $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top y \in [0, 1]\}$. $G = (V, E)$ is a Random Dot Product Graph if its adjacency matrix is drawn as $A \sim \text{BernoulliGraph}(X X^\top)$.

If on the other hand, $\mathcal{X} = \{x, y \in \mathbb{R}^{p+q} : x^\top I_{p, q} y \in [0, 1]\}$ and $A \sim \text{BernoulliGraph}(X I_{p, q} X^\top)$ for $I_{p, q} = \text{blockdiag}(I_p, -I_q)$ and $p + q = d$, then $A$ is the adjacency matrix of a Generalized Random Dot Product Graph. These are denoted by $A \sim \text{RDPG}(X)$ and $A \sim \text{GRDPG}_{p, q}(X)$ respectively.

In addition, let $F$ be a probability distribution with support $\mathcal{X}$, and $X_1, ..., X_n \stackrel{iid}{\sim} F$ with $X = \begin{bmatrix} X_1 & \cdots & X_n \end{bmatrix}^\top$. If $A$ is drawn from $X$ as before, then $(A, X) \sim \text{RDPG}(F, n)$ or $(A, X) \sim \text{GRDPG}_{p, q}(F, n)$. 
\end{definition}

@@

It is straightforward to show that all Bernoulli graphs with positive
semidefinite \(P\) are special cases of the RDPG, which is a special
case of the GRDPG, and all graphs generated by \(P\) regardless of
positive semidefiniteness are special cases of the GRDPG. This includes
the SBM, DCBM, and PABM. In the following example, we show that under
the RDPG framework, the latent configuration that induces the SBM takes
on a very particular form.

\begin{example}[Connecting the SBM to the RDPG]
Let $G = (V, E)$ with adjacency matrix $A$ be sampled from the SBM with two communities for which the within-edge probabilities are $p$ and $q$ for communities 1 and 2 respectively, and the between-community edge probability is $r$, with $p q > r^2$. Let community 1 have $n_1$ vertices, community 2 have $n_2$ vertices, and $n_1 + n_2 = n$. Without loss of generality, organize $P$ and $A$ such that the $kl^{th}$ block represents edges between communities $k$ and $l$. Then $P = \begin{bmatrix} P^{(11)} & P^{(12)} \\ P^{(21)} & P^{(22)} \end{bmatrix}$ where each block is a constant value, e.g., $P^{(11)}_{ij} = p$. Then one RDPG representation of this SBM is:

@@

We first define the PABM using the construction provided by
\citet{noroozi2019estimation}.

\begin{definition}[Popularity Adjusted Block Model]
\label{pabm}
Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
set of $n$ 
vertices, $V$. Each vertex has a community label $1, ..., K$, and the rows and 
columns of $P$ are arranged by community label such that $n_k \times n_l$ block 
$P^{(kl)}$ describes the edge probabilities between vertices in communities 
$k$ and $l$ ($P^{(lk)} = (P^{(kl)})^\top$). 
Let graph $G = (V, E)$ be an undirected, unweighted graph such 
that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
realization of $\text{Bernoulli}(P)$, i.e., 
$A_{ij} \stackrel{\text{ind}}{\sim}\text{Bernoulli}(P_{ij})$ for $i > j$ 
($A_{ij} = A_{ji}$ and $A_{ii} = 0$).

@@

It is straightforward to show that the PABM, as well as all graphs such
that the adjacency matrix is drawn such that
\(A_{ij} \stackrel{\text{ind}}{\sim}Bernoulli(P_{ij})\), is a special
case of the GRDPG. It can also be shown that the latent positions of the
PABM under the GRDPG framework consists of \(K\) \(K\)-dimensional
subspaces in \(\mathbb{R}^{K^2}\). While there is no unique latent
configuration \(X\) such that \(X X^\top = P\), the edge probability
\(P\) for the PABM, they all have this subspace structure, and one in
particular consists of \emph{orthogonal} subspaces.

@@

\hypertarget{expectation-maximization}{%
\subsubsection{Expectation
Maximization}\label{expectation-maximization}}

While direct maximization of the likelihood or log-likelihood is
NP-complete, it is possible to apply an Expectation Maximization (EM)
algorithm to this problem. To set this up, we write the complete data
log likelihood as:

\[Q = 
\sum_{i < j} \sum_{k, l} z_{ik} z_{jl} (A_{ij} \log \lambda_{il} \lambda_{jk} +
(1 - A_{ij}) \log(1 - \lambda_{il} \lambda_{jk}))
+ \sum_i \sum_k z_{ik} \log \pi_k\]

Where \(\pi_k\) are the \emph{a priori} community label probabilities.

Then the expectation step involves solving for
\(\gamma_{ik} = P(Z_i = k \mid \{\pi_l\}, \{\lambda_{jl}\})\), which is
given by

\[\log \gamma_{ik} \propto 
\log \pi_k + 
\sum_{j \neq i} \sum_l \pi_{jl} (A_{ij} \log \lambda_{il} \lambda_{jk} + 
(1 - A_{ij}) \log(1 - \lambda_{il} \lambda_{jk}))\]

For the maximization step, we can just set
\(\pi_k = \frac{1}{n} \sum_i \gamma_{ik}\), but the \(\{\lambda_{ik}\}\)
parameters cannot be maximized in closed form and require an
optimization subroutine. Our proposed work involves finding an efficient
optimization method for this subroutine.

\hypertarget{bayesian-methods}{%
\subsubsection{Bayesian Methods}\label{bayesian-methods}}

The introduction of community label probabilities \(\pi_1, ..., \pi_K\)
suggests a hierarchical model in which the community labels are
generated by a prior distribution. We can apply the same principle to
the popularity parameters \(\{\lambda_{ik}\}\) to make the model fully
Bayesian. In this setup, we sample
\(z_i \stackrel{\text{iid}}{\sim}\text{Categorical}(\pi_1, ..., \pi_K)\)
and
\(\lambda_{ik} \stackrel{\text{ind}}{\sim}\text{Beta}(a_{ik}, b_{ik})\)
before sampling the observed data
\(A_{ij} \stackrel{\text{ind}}{\sim}\lambda_{i z_j} \lambda_{j z_i}\).

The full joint distribution for this model can be written as:

\[\begin{split}
\log p & = constant \\
& + \sum_{i < j} \sum_k \sum_l z_{ik} z_{jl} 
(A_{ij} \log \lambda_{il} \lambda_{jk} + 
(1 - A_{ij}) \log (1 - \lambda_{il} \lambda_{jk})) \\
& + \sum_k \sum_i z_{ik} \log \pi_k \\
& + \sum_i \sum_k (a_{ik} - 1) \log \lambda_{ik} + (b_{ik} - 1) \log (1 - \lambda_{ik})
\end{split}\]

It is straightforward to obtain the conditional distribution for each
\(z_i\) but not each \(\lambda_{ik}\), so a Gibbs sampler cannot be
derived for this model. However, we can use a Metropolis-Hastings
algorithm. In particular, by taking a first-order Taylor approximation,
we can approximate the conditional distribution for each
\(\lambda_{ik}\) as a Beta distribution. Our proposed work involves
investigating this method by simulating this model and implementing the
Bayes sampler.

\hypertarget{generalizations-to-grdpg-based-community-detection}{%
\section{Generalizations to GRDPG-Based Community
Detection}\label{generalizations-to-grdpg-based-community-detection}}

@@

If we extend this example such that \(n_1\) points are sampled uniformly
from the segment \(f_1(t) = (t, 0)\) and \(n_2\) points are sampled
uniformly from the segment \(f_2(t) = (t, a)\) for
\(t \in [0, \cos \frac{\pi}{2} a]\), then a sample of size
\(\min(n_1, n_2) \geq N(\delta, a)\) is sufficient to satisfy:

@@

We can then further extend this to the case where points are drawn from
unit segments with noise. It is straightforward to show that if the
maximum noise is less than \(a / 3\), we can obtain a similar result as
in the noiseless case. Thus if the distribution on the two segments is
used to draw \((A, X) \sim \text{RDPG}(F, n)\), the same properties
should hold for large \(n\).

@@

Our proposed work extends these results by the following.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Show that the ASE of a (G)RDPG generated by this type of latent
  configuration produces the correct conditions as described above to
  apply the same probability statements.
\item
  Extend these results by deriving similar probability statements for
  nonuniform distributions.
\item
  Relax the condition for the minimum distance between manifolds by
  investigating alternative clustering techniques on the latent space
  (e.g., spectral clustering).
\end{enumerate}

@@

For this research, we propose using the connection between Block Models
and the (Generalized) Random Dot Product Graph to motivate community
detection methods. Previous research shows that the Stochastic Block
Model and Degree Corrected Block Model have very structured latent
configurations in the RDPG framework, and these latent structures can be
used to develop community detection algorithms that result in perfect
community detection for large \(n\). In the same vein, we investigated
the connection between the Popularity Adjusted Block Model and the GRDPG
to motivate two community detection algorithms, the first of which is an
application of an existing algorithm, Sparse Subspace Clustering, and
the second is a new algorithm based on the particular latent form of the
PABM, Orthogonal Spectral Clustering. Furthermore, we showed that for
large \(n\), our application of the SSC algorithm to the PABM upholds
the Subspace Detection Property for large yet finite \(n\), and the
output of OSC results in perfect community detection almost surely as
\(n \to \infty\). We will extend these theoretical results with
empirical simulations and real data examples. Finally, we will implement
these two community detection algorithms in an R package.

We aim to extend these results by defining more general structures in
the latent space to construct RDPG or GRDPG models and show that
community detection is possible via Adjacency Spectral Embedding. So
far, we showed that if in the latent space the communities can be
represented as points sampled uniformly on curves or one-dimensional
manifolds separated by a nonzero distance, we can apply two existing
clustering techniques to the ASE and show that they will perform perfect
community detection with high probability. We will generalize these
results to nonuniform distributions on higher dimensional manifolds as
well as investigate alternative clustering techniques for the latent
space that may perform better for smaller \(n\).
